---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Questions: 
  - Code for reading in the dataset and/or processing the data
  
```{r}
library(tidyverse)   # for the data analysis.
library(lubridate)  # for date analysis.
activity <- read.csv("activity.csv") %>%
  mutate(date = as.Date(date))
```


## What is mean total number of steps taken per day?
Questions:
 - Histogram of the total number of steps taken each day
 - Mean and median number of steps taken each day
 
```{r, summary_values, message = FALSE}
activity %>%
  group_by(date) %>%
  summarise(mean_steps = mean(steps, na.rm = TRUE), median_steps = median(steps)) %>%
  ungroup()
```

Histogram:

```{r, histogram}
activity %>%
  ggplot(aes(x = steps)) + 
  geom_histogram(binwidth = 78, fill = "red", color = "#2D2D2D") + 
  theme_light()
```


## What is the average daily activity pattern?

Questions: 
  - Time series plot of the average number of steps taken
  - The 5-minute interval that, on average, contains the maximum number of steps

Time plot series:


```{r, message = FALSE}
activity %>% 
  group_by(interval) %>%
  summarise(mean_steps = mean(steps, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(aes(x = interval, y = mean_steps)) + 
  geom_line() + 
  theme_light() + 
  labs(y = "Mean of steps", x = "Date (2012)")

```
Interval with the maximum number of steps

```{r, message = FALSE}
activity %>% 
  group_by(date) %>%
  summarise(steps = mean(steps)) %>%
  pull(steps) %>%
  max(na.rm = TRUE) -> steps_extreme

activity %>%
  group_by(date) %>%
  summarise(steps = mean(steps)) %>%
  filter(steps == steps_extreme)
# the 5-minute interval of 2012-11-23 have, on average, the maximum number of steps.
```

## Imputing missing values

Rows with missing values:


```{r}
activity %>%
  filter(is.na(steps)) %>%
  nrow()
```

To fill the missing values, we can add the mean grouped by the `interval` variable; i.e., when tha value is NA, we select all values which have the same label for identification.

```{r}
data_filled <- activity

for(index in unique(activity[["interval"]])){
  index_activity <- activity %>% 
    filter(interval == index)
  index_mean <- mean(index_activity[["steps"]], na.rm = TRUE)

  rows <- which(data_filled[["interval"]] == index)
  
  for(label_row in rows){
    if(is.na(data_filled[["steps"]][label_row])){
      data_filled[["steps"]][label_row] <- index_mean
      
    }
  }
}
```

Histograms of `data_filled`:
```{r}
data_filled %>%
  ggplot(aes(x = steps)) + 
  geom_histogram(fill = "red", color = "#2D2D2D", binwidth = 78) + 
  theme_light()
```

Mean and median of `data_filled`.

```{r, message = FALSE}
data_filled %>%
  group_by(date) %>%
  summarise(mean = mean(steps), median = median(steps)) %>%
  ungroup()
```

We hadd NaN or NA values for median and for mean before inputing the missing values; now, we have a real number in it. 

## Are there differences in activity patterns between weekdays and weekends?


```{r, message = FALSE}
data_filled %>%
  mutate(isweekend = ifelse(weekdays(date) %in% c("sábado", "domingo"), "Weekend", "Weekday")) %>% 
  group_by(interval, isweekend) %>% 
  summarise(mean_steps = mean(steps)) %>%
  ungroup() %>%
  ggplot() + 
  geom_line(aes(x = interval, y = mean_steps, group = isweekend)) + 
  theme_light() + 
  facet_grid(isweekend~.)
```
